
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(data.table)
library(glmmTMB)
library(DHARMa)
library(effects)
library(performance)
library(purrr)
library(furrr)
library(splines)
library(future.apply)
library(broom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(plotly) 
library(ggeffects)
```

# Uploading and Preparing the Data

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2013-2023.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

```{r}
# Define the 8 core intervention columns
intervention_cols <- c(
  'leading_pedestrian_interval_post', 'turn_traffic_calming_post', 
  'enhanced_crossing_post', 'speed_humps_post', 
  'street_improvement_project_post', 'street_improvement_corridors_post',
  'signal_retiming_post', 'slow_zones_post'
)

# Step 1: Find the start year for each unique intervention at each intersection.
# This creates a long-format table of every treatment start event.
intervention_start_years <- sites_ever_treated %>%
  select(intersection_id, year, all_of(intervention_cols)) %>%
  pivot_longer(
    cols = all_of(intervention_cols),
    names_to = "intervention_type",
    values_to = "is_active"
  ) %>%
  filter(is_active == 1) %>%
  group_by(intersection_id, intervention_type) %>%
  summarise(start_year = min(year), .groups = 'drop')

# Step 2: For each intersection, check if it has treatments inside vs. outside the 2015-2021 window.
# This summary is the key to aligning with your strict definition.
intersection_cohorts <- intervention_start_years %>%
  group_by(intersection_id) %>%
  summarise(
    # Does this intersection have ANY treatment that started between 2015-2021?
    treated_inside_window = any(start_year %in% 2015:2021),
    # Does this intersection have ANY treatment that started before 2015 or after 2021?
    treated_outside_window = any(start_year < 2015 | start_year > 2021)
  ) %>%
  mutate(
    # Use case_when to create the final, mutually exclusive cohort labels
    cohort = case_when(
      # If it has treatments INSIDE but NOT outside, it's "Middle Only"
      treated_inside_window & !treated_outside_window ~ "Middle Only (2015-2021)",
      
      # If it has treatments OUTSIDE but NOT inside, it's "Outside Only"
      !treated_inside_window & treated_outside_window ~ "Outside Only (2013-14 & 2022-23)",
      
      # If it has treatments in BOTH periods, it's "Mixed History"
      treated_inside_window & treated_outside_window  ~ "Mixed History",
      
      # Fallback for unexpected cases
      TRUE ~ "Error: Uncategorized"
    )
  )

# Step 3: Create the final summary table
cohort_summary_table <- intersection_cohorts %>%
  count(cohort, name = "num_intersections") %>%
  mutate(
    percentage = (num_intersections / sum(num_intersections)) * 100
  ) %>%
  arrange(desc(num_intersections))

# Final check: The total should be 21,142
total_intersections_in_r <- sum(cohort_summary_table$num_intersections)
print(paste("Total unique intersections categorized in R:", total_intersections_in_r))


# Step 4: Display the table nicely. The numbers should now match your Python audit.
knitr::kable(
  cohort_summary_table,
  caption = "Intersection Cohorts Based on Treatment Timing (Aligned with Python Validation)",
  col.names = c("Treatment Cohort", "Number of Intersections", "Percentage (%)"),
  digits = c(0, 0, 1)
)
```

# Picking a Model Type

```{r}
# fit a Poisson regression model 
# pedestrians

poisson_model <- glm(
  pedestrian_death_or_injury ~ 
    leading_pedestrian_interval_post + 
    turn_traffic_calming_post + 
    enhanced_crossing_post + 
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 +
    scaled_summonses + 
    scaled_traffic_volume +
    scaled_pop_density +
    scaled_time +
    (1 | intersection_id),
  family = "poisson", 
  data = sites_ever_treated 
)

# calculate dispersion statistic
# dispersion stat over 1.5 indicates needs for negative binomial
dispersion_stat <- sum(residuals(poisson_model, type = "pearson")^2) / poisson_model$df.residual
print(paste("Dispersion statistic:", dispersion_stat))
```

## First Model Attempt

```{r}
# pedestrians
# negative binomial regression (dispersion stat: 1.72)
# sites ever treated from 2013 through 2023
# looking at data from 2013 through 2023

model_orig <- glmmTMB(
  pedestrian_death_or_injury ~
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 +
    scaled_summonses +
    scaled_traffic_volume + 
    scaled_pop_density +
    scaled_time +
    (1 | intersection_id),
  family = nbinom1,
  data = sites_ever_treated
)

summary(model_orig)
```

```{r}
# diagnostics

# tests find zero-inflation, underdispersion, non-normal residuals, and outliers

sim_res_orig <- simulateResiduals(fittedModel = model_orig, n = 500, plot = TRUE)
testZeroInflation(sim_res_orig) # zero-inflation test
testDispersion(sim_res_orig) # dispersion test
testUniformity(sim_res_orig) # KS test
testOutliers(sim_res_orig) # outliers
```

# Fine-Tuning the Model

Will consider:
- Nested Random Effect
- Splines
- Dispersion Formula

## Nested Random Effect

A nested random effect for intersection/ NTA/ borough allows the baseline to vary by intersection, NTA, and borough.

It captures:
- Variation between regions (e.g. two boroughs), and
- Additional variation between sub-regions within the same larger region (e.g. two intersections within the same NTA).

This structure accounts for more granular clustering, recognizing that, for example: 
- Observations from the same intersection are more similar to each other than to other intersections in the same NTA.
- NTAs differ from one another, but there's also important within-NTA variation at the intersection level.
- Same applies to boroughs and NTAs, etc

```{r}
# experimenting with random effects 

# NTA and intersection nested 
model_nta_RE <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    scaled_pop_density +        
    scaled_time +       
    (1 | ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

# borough, NTA, and intersection nested 
model_boro_RE <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    scaled_pop_density +
    scaled_time +       
    (1|boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

anova(model_orig, model_nta_RE, model_boro_RE)
```

### model_boro_RE has best AIC and fixes zero-inflation and dispersion
### KS test (non-normal residuals) and outliers remain a problem for both models

```{r}
# diagnostics

sim_res_nta_RE <- simulateResiduals(fittedModel = model_nta_RE, n = 500, plot = TRUE)
testZeroInflation(sim_res_nta_RE) # zero-inflation test
testDispersion(sim_res_nta_RE) # dispersion test
testUniformity(sim_res_nta_RE) # KS test
testOutliers(sim_res_nta_RE) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_nta_RE), residuals(model_nta_RE, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

```{r}
# model 2
sim_res_boro_RE <- simulateResiduals(fittedModel = model_boro_RE, n = 500, plot = TRUE)
testZeroInflation(sim_res_boro_RE) # zero-inflation test
testDispersion(sim_res_boro_RE) # dispersion test
testUniformity(sim_res_boro_RE) # KS test
testOutliers(sim_res_boro_RE) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_boro_RE), residuals(model_boro_RE, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

## Splines

Splines allow for more flexible modeling of non-linear patterns in continuous control variables

```{r}
# check for non-linear patterns in continuous controls
# scaled_pop_density looks the least linear

ggplot(intervention_df, aes(x = scaled_summonses, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Traffic Summonses",
    x = "Scaled Traffic Summonses",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()

ggplot(intervention_df, aes(x = scaled_traffic_volume, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Traffic Volume",
    x = "Scaled Traffic Volume",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()

ggplot(intervention_df, aes(x = scaled_pop_density, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Population Density",
    x = "Scaled Population Density",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()

ggplot(intervention_df, aes(x = scaled_time, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Time",
    x = "Scaled Time",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()
```

```{r}
# updating model with spline

# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +   
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 
anova(model_boro_RE, model_spline)

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Get the confidence intervals for the log-coefficients
ci_log <- confint(model_spline, parm = "beta_", method = "wald") # Wald is fastest

# Get the coefficients table from the summary
summary_table <- summary(model_spline)$coefficients$cond

# Combine the log-estimates and log-CIs
log_results <- cbind(summary_table, ci_log)

# Exponentiate everything to get IRRs and their CIs
irr_results <- exp(log_results)

# Print the final table
print(irr_results)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  "scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  "scaled_traffic_volume",
  "scaled_summonses",
  "covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```




























































