
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(data.table)
library(glmmTMB)
library(DHARMa)
library(effects)
library(performance)
library(purrr)
library(furrr)
library(splines)
library(future.apply)
library(broom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(plotly) 
library(ggeffects)
```

# Uploading and Preparing the Data

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2013-2023.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

```{r}
# Define the 8 core intervention columns
intervention_cols <- c(
  'leading_pedestrian_interval_post', 'turn_traffic_calming_post', 
  'enhanced_crossing_post', 'speed_humps_post', 
  'street_improvement_project_post', 'street_improvement_corridors_post',
  'signal_retiming_post', 'slow_zones_post'
)

# Step 1: Find the start year for each unique intervention at each intersection.
# This creates a long-format table of every treatment start event.
intervention_start_years <- sites_ever_treated %>%
  select(intersection_id, year, all_of(intervention_cols)) %>%
  pivot_longer(
    cols = all_of(intervention_cols),
    names_to = "intervention_type",
    values_to = "is_active"
  ) %>%
  filter(is_active == 1) %>%
  group_by(intersection_id, intervention_type) %>%
  summarise(start_year = min(year), .groups = 'drop')

# Step 2: For each intersection, check if it has treatments inside vs. outside the 2015-2021 window.
# This summary is the key to aligning with your strict definition.
intersection_cohorts <- intervention_start_years %>%
  group_by(intersection_id) %>%
  summarise(
    # Does this intersection have ANY treatment that started between 2015-2021?
    treated_inside_window = any(start_year %in% 2015:2021),
    # Does this intersection have ANY treatment that started before 2015 or after 2021?
    treated_outside_window = any(start_year < 2015 | start_year > 2021)
  ) %>%
  mutate(
    # Use case_when to create the final, mutually exclusive cohort labels
    cohort = case_when(
      # If it has treatments INSIDE but NOT outside, it's "Middle Only"
      treated_inside_window & !treated_outside_window ~ "Middle Only (2015-2021)",
      
      # If it has treatments OUTSIDE but NOT inside, it's "Outside Only"
      !treated_inside_window & treated_outside_window ~ "Outside Only (2013-14 & 2022-23)",
      
      # If it has treatments in BOTH periods, it's "Mixed History"
      treated_inside_window & treated_outside_window  ~ "Mixed History",
      
      # Fallback for unexpected cases
      TRUE ~ "Error: Uncategorized"
    )
  )

# Step 3: Create the final summary table
cohort_summary_table <- intersection_cohorts %>%
  count(cohort, name = "num_intersections") %>%
  mutate(
    percentage = (num_intersections / sum(num_intersections)) * 100
  ) %>%
  arrange(desc(num_intersections))

# Final check: The total should be 21,142
total_intersections_in_r <- sum(cohort_summary_table$num_intersections)
print(paste("Total unique intersections categorized in R:", total_intersections_in_r))


# Step 4: Display the table nicely. The numbers should now match your Python audit.
knitr::kable(
  cohort_summary_table,
  caption = "Intersection Cohorts Based on Treatment Timing (Aligned with Python Validation)",
  col.names = c("Treatment Cohort", "Number of Intersections", "Percentage (%)"),
  digits = c(0, 0, 1)
)
```

```{r}
# Define the 8 core intervention columns
intervention_cols <- c(
  'leading_pedestrian_interval_post', 'turn_traffic_calming_post', 
  'enhanced_crossing_post', 'speed_humps_post', 
  'street_improvement_project_post', 'street_improvement_corridors_post',
  'signal_retiming_post', 'slow_zones_post'
)

# Step 1: Find the start year for each unique intervention at each intersection.
# (This part is unchanged and correct)
intervention_start_years <- sites_ever_treated %>%
  select(intersection_id, year, all_of(intervention_cols)) %>%
  pivot_longer(
    cols = all_of(intervention_cols),
    names_to = "intervention_type",
    values_to = "is_active"
  ) %>%
  filter(is_active == 1) %>%
  group_by(intersection_id, intervention_type) %>%
  summarise(start_year = min(year), .groups = 'drop')

# --- NEW DETAILED ANALYSIS ---

# Step 2: For each intersection, create flags for treatments in specific years/periods.
# This is the key intermediate step for our detailed breakdown.
intersection_flags <- intervention_start_years %>%
  group_by(intersection_id) %>%
  summarise(
    treated_2013 = any(start_year == 2013),
    treated_2014 = any(start_year == 2014),
    treated_2022 = any(start_year == 2022),
    treated_2023 = any(start_year == 2023),
    treated_middle = any(start_year %in% 2015:2021)
  )

# Step 3: Calculate the specific counts for each year of interest.
# NOTE: An intersection can be counted in multiple years if it had treatments in, for example, both 2013 and 2014.
summary_2013 <- intersection_flags %>% filter(treated_2013) %>% summarise(mixed = sum(treated_middle), outside_only = n() - mixed)
summary_2014 <- intersection_flags %>% filter(treated_2014) %>% summarise(mixed = sum(treated_middle), outside_only = n() - mixed)
summary_2022 <- intersection_flags %>% filter(treated_2022) %>% summarise(mixed = sum(treated_middle), outside_only = n() - mixed)
summary_2023 <- intersection_flags %>% filter(treated_2023) %>% summarise(mixed = sum(treated_middle), outside_only = n() - mixed)

# Step 4: Combine the summaries into a single long-format data frame for plotting.
plot_data <- tibble(
  year = c(2013, 2014, 2022, 2023),
  mixed_count = c(summary_2013$mixed, summary_2014$mixed, summary_2022$mixed, summary_2023$mixed),
  outside_only_count = c(summary_2013$outside_only, summary_2014$outside_only, summary_2022$outside_only, summary_2023$outside_only)
) %>%
  pivot_longer(
    cols = c(mixed_count, outside_only_count),
    names_to = "category",
    values_to = "num_intersections"
  ) %>%
  mutate(
    # Create a more descriptive factor for the plot legend
    category = factor(
      category,
      levels = c("mixed_count", "outside_only_count"),
      labels = c("Also Treated in 2015-2021 (Mixed History)", "No Treatments in 2015-2021 (Outside Only)")
    )
  )

# Step 5: Display the detailed summary table.
knitr::kable(
  plot_data %>% pivot_wider(names_from = category, values_from = num_intersections),
  caption = "Detailed Breakdown of Intersections Treated in Early/Late Years",
  col.names = c("Treatment Year", "Also Treated in 2015-2021", "Treated Outside 2015-2021 Only")
)

# Step 6: Create the informative stacked bar chart.
ggplot(plot_data, aes(x = factor(year), y = num_intersections, fill = category)) +
  geom_col(position = "stack") +
  # Add labels inside the bars
  geom_text(
    aes(label = scales::comma(num_intersections)), 
    position = position_stack(vjust = 0.5), 
    color = "white", 
    fontface = "bold",
    size = 4
  ) +
  scale_fill_brewer(palette = "Set1", name = "Intersection History") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Composition of Intersections Treated in Early (2013-14) and Late (2022-23) Years",
    subtitle = "Each bar shows the total intersections with treatments starting in that year, segmented by their history.",
    x = "Year of Treatment Start",
    y = "Number of Intersections"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size=16),
    plot.subtitle = element_text(size=11)
  )
```

# Picking a Model Type

```{r}
# fit a Poisson regression model 
# pedestrians

poisson_model <- glm(
  pedestrian_death_or_injury ~ 
    leading_pedestrian_interval_post + 
    turn_traffic_calming_post + 
    enhanced_crossing_post + 
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 +
    scaled_summonses + 
    scaled_traffic_volume +
    scaled_pop_density +
    scaled_time +
    (1 | intersection_id),
  family = "poisson", 
  data = sites_ever_treated 
)

# calculate dispersion statistic
# dispersion stat over 1.5 indicates needs for negative binomial
dispersion_stat <- sum(residuals(poisson_model, type = "pearson")^2) / poisson_model$df.residual
print(paste("Dispersion statistic:", dispersion_stat))
```

## First Model Attempt

```{r}
# pedestrians
# negative binomial regression (dispersion stat: 1.72)
# sites ever treated from 2013 through 2023
# looking at data from 2013 through 2023

model_orig <- glmmTMB(
  pedestrian_death_or_injury ~
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 +
    scaled_summonses +
    scaled_traffic_volume + 
    scaled_pop_density +
    scaled_time +
    (1 | intersection_id),
  family = nbinom1,
  data = sites_ever_treated
)

summary(model_orig)
```

```{r}
# diagnostics

# tests find zero-inflation, underdispersion, non-normal residuals, and outliers

sim_res_orig <- simulateResiduals(fittedModel = model_orig, n = 500, plot = TRUE)
testZeroInflation(sim_res_orig) # zero-inflation test
testDispersion(sim_res_orig) # dispersion test
testUniformity(sim_res_orig) # KS test
testOutliers(sim_res_orig) # outliers
```

# Fine-Tuning the Model

Will consider:
- Nested Random Effect
- Splines
- Dispersion Formula

## Nested Random Effect

A nested random effect for intersection/ NTA/ borough allows the baseline to vary by intersection, NTA, and borough.

It captures:
- Variation between regions (e.g. two boroughs), and
- Additional variation between sub-regions within the same larger region (e.g. two intersections within the same NTA).

This structure accounts for more granular clustering, recognizing that, for example: 
- Observations from the same intersection are more similar to each other than to other intersections in the same NTA.
- NTAs differ from one another, but there's also important within-NTA variation at the intersection level.
- Same applies to boroughs and NTAs, etc

```{r}
# experimenting with random effects 

# NTA and intersection nested 
model_nta_RE <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    scaled_pop_density +        
    scaled_time +       
    (1 | ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

# borough, NTA, and intersection nested 
model_boro_RE <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    scaled_pop_density +
    scaled_time +       
    (1|boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

anova(model_orig, model_nta_RE, model_boro_RE)
```

### model_boro_RE has best AIC and fixes zero-inflation and dispersion
### KS test (non-normal residuals) and outliers remain a problem for both models

```{r}
# diagnostics

sim_res_nta_RE <- simulateResiduals(fittedModel = model_nta_RE, n = 500, plot = TRUE)
testZeroInflation(sim_res_nta_RE) # zero-inflation test
testDispersion(sim_res_nta_RE) # dispersion test
testUniformity(sim_res_nta_RE) # KS test
testOutliers(sim_res_nta_RE) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_nta_RE), residuals(model_nta_RE, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

```{r}
# model 2
sim_res_boro_RE <- simulateResiduals(fittedModel = model_boro_RE, n = 500, plot = TRUE)
testZeroInflation(sim_res_boro_RE) # zero-inflation test
testDispersion(sim_res_boro_RE) # dispersion test
testUniformity(sim_res_boro_RE) # KS test
testOutliers(sim_res_boro_RE) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_boro_RE), residuals(model_boro_RE, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

## Splines

Splines allow for more flexible modeling of non-linear patterns in continuous control variables

```{r}
# check for non-linear patterns in continuous controls
# scaled_pop_density looks the least linear

ggplot(intervention_df, aes(x = scaled_summonses, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Traffic Summonses",
    x = "Scaled Traffic Summonses",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()

ggplot(intervention_df, aes(x = scaled_traffic_volume, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Traffic Volume",
    x = "Scaled Traffic Volume",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()

ggplot(intervention_df, aes(x = scaled_pop_density, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Population Density",
    x = "Scaled Population Density",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()

ggplot(intervention_df, aes(x = scaled_time, y = pedestrian_death_or_injury)) +
  geom_point(alpha = 0.05, shape = 16) + 
  labs(
    title = "Pedestrian Deaths/Injuries vs. Scaled Time",
    x = "Scaled Time",
    y = "Pedestrian Deaths/Injuries"
  ) +
  theme_minimal()
```

```{r}
# updating model with spline

# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +   
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 
anova(model_boro_RE, model_spline)

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Get the confidence intervals for the log-coefficients
ci_log <- confint(model_spline, parm = "beta_", method = "wald") # Wald is fastest

# Get the coefficients table from the summary
summary_table <- summary(model_spline)$coefficients$cond

# Combine the log-estimates and log-CIs
log_results <- cbind(summary_table, ci_log)

# Exponentiate everything to get IRRs and their CIs
irr_results <- exp(log_results)

# Print the final table
print(irr_results)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  "scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  "scaled_traffic_volume",
  "scaled_summonses",
  "covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# compare to scatter plots/ use judgement when thinking about whether or not the splines capture the trend

# plotting the relationship captured by each model
plot(Effect("scaled_pop_density", model_orig))
plot(Effect("scaled_pop_density", model_spline))
```

```{r}
# after introducing new terms, is there multicollinearity?

# no issues

check_collinearity(model_spline)
```

### model_spline has improved AIC and unchanged diagnostics, so will choose it

```{r}
# diagnostics

sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_spline), residuals(model_spline, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

# Dispersion Formula (dispformula)

Conditional model:
- The mean (expected outcome): How predictors influence the average pedestrian casualties per unit

Dispersion model:	
- The variance (spread of outcome): How predictors influence the dispersion — i.e., whether the variance is larger or smaller for certain values

Without dispformula, the model assumes the same dispersion level across all observations. By adding dispformula, it explicitly models heteroskedasticity — i.e., allowing areas or time periods to have more or less stable outcomes.

This can lead to better residuals, more accurate standard errors, and more trustworthy p-values.

```{r}
# timing for how long one simulation will take 

timing <- system.time({
  
mod <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +   
    (1 | boro_name/ntaname/intersection_id),
  dispformula = ~1,
  family = nbinom1,
  data = sites_ever_treated
) 
  
sim_res <- simulateResiduals(mod)
dharma_disp <- testDispersion(sim_res, plot = FALSE)
dispersion_numeric <- as.numeric(dharma_disp$statistic)
    
})

print(timing)
```

```{r}
# using a simulation to find best dispersion formula for the NB model
# considering AIC, BIC, and KS test (normality of residuals, which is still failing the DHARMa test)

# seconds to complete a model test (estimate based on running one test)
secs_to_complete_one = timing["elapsed"]

# number of cores working at once
num_workers = 5
plan(multisession, workers = num_workers)

# formula to test with
fixed_formula <- pedestrian_death_or_injury ~ 
  turn_traffic_calming_post +
  leading_pedestrian_interval_post +
  enhanced_crossing_post +
  speed_humps_post +
  street_improvement_project_post +
  street_improvement_corridors_post +
  signal_retiming_post +
  slow_zones_post +
  speed_limit_post +
  covid_2020 + 
  scaled_traffic_volume + 
  splines::ns(scaled_pop_density, df = 2) +
  scaled_time +
  (1 | boro_name/ntaname/intersection_id)

# list of possible variables for dispformula (just doing controls)
disp_vars <- c(
  "scaled_time",
  "scaled_pop_density",
  "scaled_traffic_volume",
  "scaled_summonses",
  "covid_2020",
  "speed_limit_post"
  # "slow_zones_post",
  # "signal_retiming_post",
  # "street_improvement_corridors_post",
  # "street_improvement_project_post",
  # "speed_humps_post",
  # "enhanced_crossing_post",
  # "leading_pedestrian_interval_post",
  # "turn_traffic_calming_post"
)

# set number of variables in dispformula 
max_terms <- 5
# calculate number of possible combinations
disp_combinations <- unlist(
  lapply(0:min(max_terms, length(disp_vars)), function(i) combn(disp_vars, i, simplify = FALSE)),
  recursive = FALSE
)

cat("Number of dispersion combinations:", length(disp_combinations), "\n")
cat("Estimated completion time:", length(disp_combinations) * secs_to_complete_one /  num_workers / 60 / 60, "hours\n")

# run model with one dispformula combo and collect stats
fit_model_safe <- function(disp_vars_subset) {
  disp_formula_str <- if (length(disp_vars_subset) == 0) {
    "~ 1"
  } else {
    paste0("~ ", paste(disp_vars_subset, collapse = " + "))
  }

  disp_formula <- as.formula(disp_formula_str)

  tryCatch({
    mod <- glmmTMB(
      formula = fixed_formula,
      dispformula = disp_formula,
      family = nbinom1,
      data = sites_ever_treated
    )
    
    # DHARMa KS test for uniformity
    simulation_output <- simulateResiduals(fittedModel = mod)
    dharma_ks <- tryCatch({
      testUniformity(simulation_output, plot = FALSE)
    }, error = function(e) {
      list(statistic = NA, p.value = NA)
    })

    # extracting statistic
    ks_numeric <- as.numeric(dharma_ks$statistic)
     
    # collecting values for AIC, BIC, and KS stat   
    list(
      dispformula = disp_formula_str,
      AIC = AIC(mod),
      BIC = BIC(mod),
      dharma_ks_stat = ks_numeric,
      dharma_ks_p = dharma_ks$p.value,
      error = NA
    ) 
  }, error = function(e) { # noting any errors 
    list(
      dispformula = disp_formula_str,
      AIC = NA,
      BIC = NA,
      dharma_ks_stat = NA,
      dharma_ks_p = NA,
      error = e$message
    )
  })
}

# run in parallel with future_map
results_disp <- future_map(
  disp_combinations,
  fit_model_safe,
  .progress = TRUE,
  .options = furrr_options(seed = TRUE)
)

# convert to dataframe
results_df_disp <- bind_rows(results_disp)
results_df_disp %>% count(error, sort = TRUE)

# show best model for AIC
best_model_row_AIC <- results_df_disp %>% filter(!is.na(AIC)) %>% slice_min(AIC, n = 1)
print(best_model_row_AIC)

# show best model for BIC
best_model_row_BIC <- results_df_disp %>% filter(!is.na(BIC)) %>% slice_min(BIC, n = 1)
print(best_model_row_BIC)

# show best model for KS (highest p-val)
best_model_row_KS <- results_df_disp %>% filter(!is.na(dharma_ks_p)) %>% slice_max(dharma_ks_p, n = 1)
print(best_model_row_KS)

# show best combination of AIC, BIC, and KS (by combining their ranks)
# lowest combined score is the best
results_df_disp <- results_df_disp %>%
  filter(!is.na(AIC), !is.na(BIC), !is.na(dharma_ks_p)) %>%
  mutate(
    aic_rank = rank(AIC),
    bic_rank = rank(BIC),
    ks_score = 1 - dharma_ks_p, # higher p-val is better
    combined_rank = aic_rank + bic_rank + ks_score
    # optionally apply weights, e.g., combined_rank = aic_rank + bic_rank + 10 * ks_score
  ) %>%
  arrange(combined_rank)

# view the top models
head(results_df_disp, 5)

```

```{r}
# best overall ~ scaled_traffic_volume + scaled_summonses + covid_2020
# best for KS test ~ scaled_pop_density + covid_2020

# testing both

model_dispformula1 <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +       
    (1 | boro_name/ntaname/intersection_id),
  dispformula = ~ scaled_traffic_volume + scaled_summonses + covid_2020,
  family = nbinom1,
  data = sites_ever_treated
) 

model_dispformula2 <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +       
    (1 | boro_name/ntaname/intersection_id),
  dispformula = ~ scaled_pop_density + covid_2020,
  family = nbinom1,
  data = sites_ever_treated
) 

anova(model_spline, model_dispformula1, model_dispformula2)
```

### model_dispformula1 has slightly better AIC but no improvement on diagnostics, so will stick with previous model

```{r}
# diagnostics

sim_res_dispformula1 <- simulateResiduals(fittedModel = model_dispformula1, n = 500, plot = TRUE)
testZeroInflation(sim_res_dispformula1) # zero-inflation test
testDispersion(sim_res_dispformula1) # dispersion test
testUniformity(sim_res_dispformula1) # KS test
testOutliers(sim_res_dispformula1) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_dispformula1), residuals(model_dispformula1, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

```{r}
# model 2
sim_res_dispformula2 <- simulateResiduals(fittedModel = model_dispformula2, n = 500, plot = TRUE)
testZeroInflation(sim_res_dispformula2) # zero-inflation test
testDispersion(sim_res_dispformula2) # dispersion test
testUniformity(sim_res_dispformula2) # KS test
testOutliers(sim_res_dispformula2) # outliers
```

```{r}
# plotting residuals for model
# exhibits hetereoskedasticity

plot(fitted(model_dispformula2), residuals(model_dispformula2, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals",
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")
```

# Exploring the Chosen Model

```{r}
# looking at final diagnostic plots for model (greater number of simulations, plus in-depth outlier check)

sim_res <- simulateResiduals(fittedModel = model_spline, n = 1000, plot = TRUE)
testZeroInflation(sim_res) # zero-inflation test
testDispersion(sim_res) # dispersion test
testUniformity(sim_res) # KS test
# DHARMa:testOutliers with type = binomial may have inflated Type I error rates for integer-valued distributions. To get a more exact result, it is recommended to re-run testOutliers with type = 'bootstrap'. See ?testOutliers for details
testOutliers(sim_res, type = "bootstrap") # outlier test
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  "scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  "scaled_traffic_volume",
  "scaled_summonses",
  "covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# example marginal (isolated) effects of interventions

# turn traffic calming
p1 <- ggpredict(model_spline, terms = "turn_traffic_calming_post [all]") %>%
  plot() + 
  ggtitle("Effect of Turn Traffic Calming")

p1

# speed humps
p2 <- ggpredict(model_spline, terms = "speed_humps_post [all]") %>%
  plot() + 
  ggtitle("Effect of Speed Humps")

p2

# signal retiming
p3 <- ggpredict(model_spline, terms = "signal_retiming_post [all]") %>%
  plot() + 
  ggtitle("Effect of Signal Retiming")

p3
```
































