
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

This document provides the additional analyses and diagnostic checks for the manuscript. It includes:

1.  A visualization of the overall trend in pedestrian casualties in the analytic sample.
2.  A series of visual and formal tests of the parallel trends assumption, a critical requirement for the staggered difference-in-difference design.

All analyses are conducted using the final analytic dataset as described in the main paper.

## 1. Data Loading and Preparation

```{r data_load_and_prep}
# Load all required libraries for the entire analysis
library(tidyverse)
library(data.table)
library(glmmTMB)
library(splines)
library(broom)
library(broom.mixed)
library(ggplot2)
library(tools) 
library(forcats)
library(patchwork)
library(kableExtra) 
library(Kendall)

# Set a consistent global theme for all plots
theme_set(theme_minimal(base_size = 14))

# Load the main analytic dataset
file_path <- '../data/output/intersection_intervention_table_ever_treated_2015-2022.csv'

if (!file.exists(file_path)) {
  stop("Data file not found. Please check the path in the 'file_path' variable.")
}

sites_ever_treated <- read_csv(file_path, show_col_types = FALSE)

# Scale continuous variables for use in regression models
sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )

print("Data loaded and prepared successfully.")
```

---

## 2. Overall Trend in Pedestrian Casualties (Figure 1)

This section generates Figure 1, which displays the total number of annual pedestrian casualties within our analytic sample from 2013 to 2023. A Mann-Kendall test is performed to formally test for a monotonic trend.

```{r figure_1_total_casualties}
# Summarize the data by year, calculating the sum of casualties.
yearly_totals <- sites_ever_treated %>%
  group_by(year) %>%
  summarise(
    total_casualties = sum(pedestrian_death_or_injury, na.rm = TRUE),
    .groups = 'drop'
  )

# Perform a Mann-Kendall trend test on the time series of total casualties.
mk_test_result_total <- MannKendall(yearly_totals$total_casualties)

# Format the p-value for display on the plot.
p_value_total <- mk_test_result_total$sl
if (p_value_total < 0.001) {
  p_text_total <- "Mann-Kendall p < 0.001"
} else {
  p_text_total <- sprintf("Mann-Kendall p = %.3f", p_value_total)
}

# Generate the plot.
ggplot(yearly_totals, aes(x = year, y = total_casualties)) +
  geom_line(linewidth = 1.2, color = "darkred") +
  geom_point(size = 3, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_vline(xintercept = 2014.5, linetype = "dashed", color = "darkgreen", alpha = 0.7) +
  annotate("text", x = 2014.5, y = max(yearly_totals$total_casualties), 
           label = "Vision Zero Begins", hjust = -0.1, size = 3.5, color = "darkgreen") +
  annotate("text", x = max(yearly_totals$year), y = max(yearly_totals$total_casualties), 
           label = p_text_total, hjust = 1.1, vjust = 1.5, size = 4, color = "black", fontface = "italic") +
  scale_x_continuous(breaks = seq(min(yearly_totals$year), max(yearly_totals$year), by = 1)) +
  labs(
    title = "Total Annual Pedestrian Casualties in the Analytic Sample",
    subtitle = "Showing the unadjusted citywide trend from 2013 to 2023",
    x = "Year",
    y = "Total Annual Pedestrian Casualties"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Load the main analytic dataset
file_path <- '../data/output/intersection_intervention_table_final.csv'

if (!file.exists(file_path)) {
  stop("Data file not found. Please check the path in the 'file_path' variable.")
}

full_data <- read_csv(file_path, show_col_types = FALSE)

# Summarize the data by year, calculating the sum of casualties.
yearly_totals <- full_data %>%
  group_by(year) %>%
  summarise(
    total_casualties = sum(pedestrian_death_or_injury, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  filter(year != 2024)

# Perform a Mann-Kendall trend test on the time series of total casualties.
mk_test_result_total <- MannKendall(yearly_totals$total_casualties)

# Format the p-value for display on the plot.
p_value_total <- mk_test_result_total$sl
if (p_value_total < 0.001) {
  p_text_total <- "Mann-Kendall p < 0.001"
} else {
  p_text_total <- sprintf("Mann-Kendall p = %.3f", p_value_total)
}

# Generate the plot.
ggplot(yearly_totals, aes(x = year, y = total_casualties)) +
  geom_line(linewidth = 1.2, color = "darkred") +
  geom_point(size = 3, color = "darkred") +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_vline(xintercept = 2014.5, linetype = "dashed", color = "darkgreen", alpha = 0.7) +
  annotate("text", x = 2014.5, y = max(yearly_totals$total_casualties), 
           label = "Vision Zero Begins", hjust = -0.1, size = 3.5, color = "darkgreen") +
  annotate("text", x = max(yearly_totals$year), y = max(yearly_totals$total_casualties), 
           label = p_text_total, hjust = 1.1, vjust = 1.5, size = 4, color = "black", fontface = "italic") +
  scale_x_continuous(breaks = seq(min(yearly_totals$year), max(yearly_totals$year), by = 1)) +
  labs(
    title = "Total Annual Pedestrian Casualties in NYC",
    subtitle = "Showing the unadjusted citywide trend from 2013 to 2023",
    x = "Year",
    y = "Total Annual Pedestrian Casualties"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

## 3. Parallel Trends Assumption Checks

This section contains the three diagnostic checks for the parallel trends assumption.

### 3.1 Visual Check by Treatment Cohort (Figure 2)

This plot provides a descriptive check of the assumption. It groups intersections by their first treatment year and plots their average raw casualty trends over calendar time. Parallel movement of the lines prior to their respective treatment start dates provides visual support for the assumption.

```{r figure_2_cohort_plot}
# Define the non-universal treatment variables.
treatment_vars <- c(
  "turn_traffic_calming_post", "leading_pedestrian_interval_post", 
  "enhanced_crossing_post", "speed_humps_post", "street_improvement_project_post", 
  "street_improvement_corridors_post", "signal_retiming_post", "slow_zones_post"
)

# Identify the first treatment year for each intersection.
first_treatment_year_df <- sites_ever_treated %>%
  select(intersection_id, year, all_of(treatment_vars)) %>%
  pivot_longer(
    cols = all_of(treatment_vars),
    names_to = "intervention",
    values_to = "is_treated"
  ) %>%
  filter(is_treated == 1) %>%
  group_by(intersection_id) %>%
  summarise(first_treat_year = min(year), .groups = 'drop')

# Assign intersections to cohorts and set a specific factor order for the legend.
sites_with_cohorts <- sites_ever_treated %>%
  left_join(first_treatment_year_df, by = "intersection_id") %>%
  mutate(
    cohort = case_when(
      first_treat_year %in% c(2015, 2016) ~ "Early Adopters (2015-16)",
      first_treat_year %in% c(2017, 2018, 2019) ~ "Middle Adopters (2017-19)",
      first_treat_year %in% c(2020, 2021) ~ "Late Adopters (2020-21)",
      TRUE ~ NA_character_
    ),
    cohort = factor(cohort, levels = c("Early Adopters (2015-16)", "Middle Adopters (2017-19)", "Late Adopters (2020-21)"))
  ) %>%
  filter(!is.na(cohort))

# Calculate the mean casualties per cohort for each year.
cohort_trends <- sites_with_cohorts %>%
  group_by(cohort, year) %>%
  summarise(
    mean_casualties = mean(pedestrian_death_or_injury, na.rm = TRUE),
    .groups = 'drop'
  )

# Define a consistent color palette for the cohorts.
cohort_colors <- c(
  "Early Adopters (2015-16)" = "#00BA38", # Green
  "Middle Adopters (2017-19)" = "#F8766D", # Red
  "Late Adopters (2020-21)" = "#619CFF"  # Blue
)

# Create a helper data frame for the vertical line legends.
vline_data <- tibble(
  xintercept = c(2015, 2017, 2020),
  cohort = factor(c("Early Adopters (2015-16)", "Middle Adopters (2017-19)", "Late Adopters (2020-21)"),
                  levels = c("Early Adopters (2015-16)", "Middle Adopters (2017-19)", "Late Adopters (2020-21)")),
  legend_label = factor(c("Treatment Starts (Early)", "Treatment Starts (Middle)", "Treatment Starts (Late)"),
                        levels = c("Treatment Starts (Early)", "Treatment Starts (Middle)", "Treatment Starts (Late)"))
)

# Generate the plot.
ggplot(data = cohort_trends) +
  geom_line(aes(x = year, y = mean_casualties, color = cohort, group = cohort), linewidth = 1.2) +
  geom_point(aes(x = year, y = mean_casualties, color = cohort, group = cohort), size = 2.5) +
  geom_vline(
    data = vline_data,
    aes(xintercept = xintercept, color = cohort, linetype = legend_label),
    alpha = 0.8
  ) +
  scale_color_manual(name = "Treatment Cohort", values = cohort_colors) +
  scale_linetype_manual(name = "Event", values = rep("dashed", 3)) +
  guides(
    color = guide_legend(order = 1, title = "Treatment Cohort"),
    linetype = guide_legend(
      order = 2, 
      title = "Event",
      override.aes = list(color = cohort_colors[levels(vline_data$cohort)])
    )
  ) +
  scale_x_continuous(breaks = seq(min(cohort_trends$year), max(cohort_trends$year), by = 1)) +
  labs(
    title = "Average Pedestrian Casualties by Treatment Adoption Cohort",
    subtitle = "A Visual Check for the Parallel Trends Assumption",
    x = "Year",
    y = "Average Annual Casualties per Intersection"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.box = "vertical"
  )
```

### 3.2 Formal Test 1: Global Event Study (Figure 3)

This plot shows the results of a formal event study that aligns all intersections by their first treatment year (`t=0`). It provides a statistical test for differential pre-trends for the average treated intersection in the sample. The non-significance of the pre-treatment coefficients supports the parallel trends assumption.

```{r figure_3_global_event_study}
# Align data in event time relative to the first treatment year.
sites_with_event_year <- sites_ever_treated %>%
  left_join(first_treatment_year_df, by = "intersection_id") %>%
  mutate(
    time_to_treatment = year - first_treat_year,
    time_to_treatment_binned = case_when(
      is.na(time_to_treatment) ~ NA_real_,
      time_to_treatment <= -3 ~ -3,
      time_to_treatment >= 3  ~ 3,
      TRUE ~ time_to_treatment
    ),
    time_to_treatment_factor = factor(time_to_treatment_binned)
  )

# Set t-1 as the reference category.
sites_with_event_year$time_to_treatment_factor <- relevel(sites_with_event_year$time_to_treatment_factor, ref = "-1")

# Estimate the global event study model.
global_event_study_model <- glmmTMB(
  pedestrian_death_or_injury ~ 
    time_to_treatment_factor +
    speed_limit_post + 
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_with_event_year,
  na.action = na.omit # Ensure intersections without a first_treat_year are omitted
) 

# Extract coefficients for plotting.
event_study_results <- broom.mixed::tidy(global_event_study_model, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(grepl("time_to_treatment_factor", term)) %>%
  mutate(
    relative_time = as.numeric(gsub("time_to_treatment_factor", "", term))
  )

reference_period <- tibble(
  term = "t = -1", estimate = 1, conf.low = 1, conf.high = 1, relative_time = -1
)

plot_data <- bind_rows(event_study_results, reference_period) %>%
  arrange(relative_time)

# Create a helper data frame for annotations.
annotation_df <- tibble(
  x = c(-2, 1.5),
  y = max(plot_data$conf.high, na.rm = TRUE) * 1.05, 
  label = c("Pre-Treatment", "Post-Treatment"),
  color = c("red4", "blue4")
)

# Generate the plot.
ggplot(plot_data, aes(x = relative_time, y = estimate)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = -0.5, linetype = "dotted", color = "red") +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.8) +
  geom_text(
    data = annotation_df,
    aes(x = x, y = y, label = label),
    color = annotation_df$color,
    fontface = "bold",
    size = 4.5
  ) +
  labs(
    title = "Global Event Study: Effect of First VZ Intervention",
    subtitle = "Relative to 1 year prior to treatment (t = -1)",
    x = "Years Relative to First Intervention",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  scale_x_continuous(breaks = min(plot_data$relative_time):max(plot_data$relative_time)) +
  coord_cartesian(ylim = c(NA, max(plot_data$conf.high, na.rm = TRUE) * 1.1))
```

### 3.3 Formal Test 2: Intervention-Specific Event Studies (Figure 4)

This is our most robust check. It runs nine separate event studies, one for each intervention. Each model uses a "pure" control group of intersections that never received the target treatment. The results are combined into a single multi-panel figure for a comprehensive overview.

```{r figure_4_multi_panel_event_study}
# Define the function to create a single, panel-friendly plot.
run_event_study_for_panel <- function(data, target_intervention) {
  all_interventions <- c(
    "turn_traffic_calming_post", "leading_pedestrian_interval_post", 
    "enhanced_crossing_post", "speed_humps_post", "street_improvement_project_post", 
    "street_improvement_corridors_post", "signal_retiming_post", "slow_zones_post",
    "speed_limit_post" 
  )
  control_interventions <- setdiff(all_interventions, target_intervention)
  event_year_df <- data %>%
    filter(.data[[target_intervention]] == 1) %>%
    group_by(intersection_id) %>%
    summarise(event_year = min(year), .groups = 'drop')
  if (nrow(event_year_df) == 0) return(NULL)
  analysis_data <- data %>%
    left_join(event_year_df, by = "intersection_id") %>%
    mutate(
      time_to_treatment = year - event_year,
      time_to_treatment_binned = case_when(
        is.na(time_to_treatment) ~ NA_real_,
        time_to_treatment <= -3 ~ -3,
        time_to_treatment >= 3  ~ 3,
        TRUE ~ time_to_treatment
      ),
      time_to_treatment_factor = factor(time_to_treatment_binned),
      time_to_treatment_factor = forcats::fct_explicit_na(time_to_treatment_factor, na_level = "Control")
    )
  if (any(analysis_data$time_to_treatment_factor != "Control")) {
    analysis_data$time_to_treatment_factor <- relevel(analysis_data$time_to_treatment_factor, ref = "-1")
  } else {
    return(NULL)
  }
  model_formula <- as.formula(
    paste(
      "pedestrian_death_or_injury ~ time_to_treatment_factor +",
      paste(control_interventions, collapse = " + "),
      "+ covid_2020 + scaled_summonses + scaled_traffic_volume + ns(scaled_pop_density, df = 2)",
      "+ (1 | boro_name/ntaname/intersection_id)"
    )
  )
  event_model <- tryCatch({
    glmmTMB(model_formula, family = nbinom1, data = analysis_data)
  }, error = function(e) {
    message(paste("Model for", target_intervention, "failed to converge:", e$message))
    return(NULL)
  })
  if (is.null(event_model)) return(NULL)
  results <- broom.mixed::tidy(event_model, conf.int = TRUE, exponentiate = TRUE) %>%
    filter(grepl("time_to_treatment_factor", term) & !grepl("Control", term)) %>%
    mutate(
      relative_time = as.numeric(gsub("time_to_treatment_factor", "", term))
    )
  reference_period <- tibble(
    term = "t = -1", estimate = 1, conf.low = 1, conf.high = 1, relative_time = -1
  )
  plot_data <- bind_rows(results, reference_period) %>%
    arrange(relative_time)
  plot_title <- toTitleCase(gsub("_post", "", target_intervention))
  plot_title <- gsub("_", " ", plot_title)
  
  p <- ggplot(plot_data, aes(x = relative_time, y = estimate)) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
    geom_vline(xintercept = -0.5, linetype = "dotted", color = "red") +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.5) +
    labs(
      title = plot_title, 
      x = NULL,
      y = NULL
    ) +
    scale_x_continuous(breaks = seq(min(plot_data$relative_time, na.rm=TRUE), max(plot_data$relative_time, na.rm=TRUE), by=2)) +
    theme_minimal(base_size = 9)
    
  return(p)
}

# Define the list of interventions to loop through.
interventions_to_test <- c(
  "slow_zones_post",
  "turn_traffic_calming_post",
  "speed_humps_post",
  "street_improvement_corridors_post",
  "signal_retiming_post",
  "leading_pedestrian_interval_post",
  "enhanced_crossing_post",
  "street_improvement_project_post",
  "speed_limit_post"
)

# Initialize an empty list to store the plots.
panel_plots <- list()

# Loop through each intervention, run the model, and store the resulting plot.
cat("--- Running intervention-specific event studies and collecting plots... ---\n")
for (intervention in interventions_to_test) {
  cat("  - Running for:", intervention, "\n")
  panel_plots[[intervention]] <- run_event_study_for_panel(sites_ever_treated, intervention)
}
cat("--- All plots collected. ---\n")


# Assemble the final multi-panel figure using the patchwork package.
multi_panel_figure <- wrap_plots(panel_plots, ncol = 3)

final_figure <- multi_panel_figure +
  plot_annotation(
    title = "Intervention-Specific Event Study Analyses",
    subtitle = "All estimates are relative to 1 year prior to treatment (t = -1)",
    tag_levels = 'a' 
  ) &
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 11),
    plot.tag = element_text(size = 8, face = "bold")
  )

# Add common axis labels to the combined figure.
final_figure_with_axes <- wrap_elements(panel = final_figure) +
  labs(x = "Years Relative to Intervention", y = "Incidence Rate Ratio (IRR)") +
  theme(
    plot.background = element_rect(fill = "white", color = NA)
  )

# Display the final, combined figure.
print(final_figure_with_axes)
```














