
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(data.table)
library(glmmTMB)
library(DHARMa)
library(effects)
library(performance)
library(purrr)
library(furrr)
library(splines)
library(future.apply)
library(broom)
library(broom.mixed)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(plotly) 
library(ggeffects)
```

# Uploading and Preparing the Data

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

# Model - Year as a categorical variable

```{r}
# convert year to categorical
sites_ever_treated$year <- as.factor(sites_ever_treated$year)

# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    year +   
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Tidy the model output. This will automatically handle the NA coefficients.
irr_results_tidy <- tidy(model_spline, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         effects = "fixed") %>% # Only show fixed effects
                   select(term, estimate, conf.low, conf.high, p.value) # Optional: reorder

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# after introducing new term, is there multicollinearity?

# very high

check_collinearity(model_spline)
```

```{r}
# diagnostics
# cannot do because of NA coefficients 
sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
```

# There is a big change in speed limit post (significantly negative -> significantly positive), slow zones (significantly negative -> non significant), and signal retiming (significantly positive -> non significant).











# Model - Year as a categorical variable, taking out all other time dependent effects

```{r}
# convert year to categorical
sites_ever_treated$year <- as.factor(sites_ever_treated$year)

# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    #covid_2020 + 
    #scaled_summonses + 
    #scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    year +   
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Tidy the model output. This will automatically handle the NA coefficients.
irr_results_tidy <- tidy(model_spline, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         effects = "fixed") %>% # Only show fixed effects
                   select(term, estimate, conf.low, conf.high, p.value) # Optional: reorder

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# after removing terms, is there multicollinearity?

# very high for speed limit - this is probably what's causing the issue

check_collinearity(model_spline)
```

```{r}
# diagnostics
sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
AIC(model_spline)
BIC(model_spline)
```

# No imporvement for any diagnostic measure, if anythinng some got even more significant
# Worse AIC and BIC too

# There is a big change in speed limit post (significantly negative -> significantly positive), slow zones (significantly negative -> non significant), and signal retiming (significantly positive -> non significant).













# Model - Poisson version, all else held equal

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

```{r}
# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +   
    (1 | boro_name/ntaname/intersection_id),
  family = poisson,
  data = sites_ever_treated
) 

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Tidy the model output. This will automatically handle the NA coefficients.
irr_results_tidy <- tidy(model_spline, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         effects = "fixed") %>% # Only show fixed effects
                   select(term, estimate, conf.low, conf.high, p.value) # Optional: reorder

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# is there multicollinearity?

# no issues

check_collinearity(model_spline)
```

```{r}
# diagnostics

sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
AIC(model_spline)
BIC(model_spline)
```

# Same exact results










# Model - Linear version, all else held equal

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

```{r}
# one spline for scaled_pop_density
model_linear <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +   
    (1 | boro_name/ntaname/intersection_id),
  family = gaussian,
  data = sites_ever_treated
) 

summary(model_linear)
```

```{r}
# visualize the linear model

library(plotly)
library(dplyr)
library(glmmTMB)

# 1. Run the Linear Model (if not already in your environment)
# model_linear <- glmmTMB(
#   pedestrian_death_or_injury ~ 
#     turn_traffic_calming_post +
#     leading_pedestrian_interval_post +
#     enhanced_crossing_post +
#     speed_humps_post +
#     street_improvement_project_post +
#     street_improvement_corridors_post +
#     signal_retiming_post +
#     slow_zones_post +
#     speed_limit_post +
#     covid_2020 + 
#     scaled_summonses + 
#     scaled_traffic_volume + 
#     ns(scaled_pop_density, df = 2) +        
#     scaled_time +   
#     (1 | boro_name/ntaname/intersection_id),
#   family = gaussian(),
#   data = sites_ever_treated
# )

# 2. Extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_linear)$cond
model_vcov <- vcov(model_linear)$cond
model_se <- sqrt(diag(model_vcov))

# 3. Compute z-scores and p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# 4. Compute confidence intervals (Raw Additive Scale)
# No exponentiation here
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# 5. Define custom order
custom_order <- c(
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# 6. Create plotting data frame
# Using raw estimates instead of percent change
plot_df <- data.frame(
  term = names(coef_point_estimates),
  estimate = coef_point_estimates,
  lower_ci = ci_lower,
  upper_ci = ci_upper,
  p_value = p_values_model
) %>%
  # Filter to only include the specific terms of interest
  filter(term %in% custom_order) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# 7. Create Interactive Plotly Chart
p <- plot_ly(data = plot_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~estimate,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "<b>Term:</b> ", term, "<br>",
      "<b>Estimate:</b> ", round(estimate, 4), "<br>",
      "<b>95% CI:</b> (", round(lower_ci, 4), ", ", round(upper_ci, 4), ")<br>",
      "<b>p-value:</b> ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries (Linear Model)",
    subtitle = "Additive Fixed Effects (Gaussian family)",
    xaxis = list(
      title = "Additive Effect on Outcome", 
      zeroline = TRUE, 
      zerolinecolor = 'black', # The null line at 0
      zerolinewidth = 2
    ),
    yaxis = list(
      title = "", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red"),
    margin = list(l = 150) # Margin for variable names
  )

# Show the plot
p
```

```{r}
# is there multicollinearity?

# no issues

check_collinearity(model_linear)
```

# Some coefficients lose or gain significane (speed humps, SIP corridors, and signal retiming become non-significant, LPI becomes significant and negative), but no reversals from significant increase/decrease to singificant decrease/increase.
# Some inconsistencies are to be expected due to differences between model types.














# Model - Random effects as fixed instead

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

```{r}
# Install if needed
if (!require("fixest")) install.packages("fixest")
library(fixest)

# Run Fixed Effects Negative Binomial
model_fixed <- fenegbin(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    # Note: remove boro and nta because they are invariant within an intersection
    ns(scaled_pop_density, df = 2) +        
    scaled_time | 
    intersection_id,
  data = sites_ever_treated
)

summary(model_fixed)
```

```{r}
# exploring model_spline since it is the final chosen model

# Load broom (fixest works with broom automatically)
library(broom)

# Tidy the model output
# fixest automatically clusters SEs by the fixed effect (intersection_id)
irr_results_tidy <- tidy(model_fixed, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         se = "cluster") %>%   # Explicitly request clustered SEs
                   select(term, estimate, conf.low, conf.high, p.value)

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results
library(plotly)
library(dplyr)
library(fixest)

# --- FIX: Extraction method ---
# We extract from the summary table to ensure row counts match (ignoring theta)
# se = "cluster" ensures we use the clustered SEs you requested
model_summary <- summary(model_fixed, se = "cluster")$coeftable

# Extract columns by index to be safe (1=Est, 2=SE, 4=Pval)
coef_point_estimates <- model_summary[, 1]
model_se <- model_summary[, 2]
p_values_model <- model_summary[, 4]

# compute z-scores (Optional, since p-values are already extracted above, but good for checking)
z_scores <- coef_point_estimates / model_se

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", 
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
# Using names(coef_point_estimates) works now because we extracted from the aligned table
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  # Filter to only include terms in the custom order
  filter(term %in% custom_order) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Fixed Effects Negative Binomial; SEs clustered by Intersection",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p
```

# Speed humps and SIP corridors go from negative and significant to negative and non-significant. Signal retiming goes from positive and significant to non-significant.  LPI goes from non significant to negative and significant. 















# Model - Random effects as fixed instead, year as categorical (have to remove all other covariates, else no CIs can be calculated)

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )

# convert year to categorical
sites_ever_treated$year <- as.factor(sites_ever_treated$year)
```

```{r}
# Install if needed
if (!require("fixest")) install.packages("fixest")
library(fixest)

# Run Fixed Effects Negative Binomial
model_fixed <- fenegbin(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    #covid_2020 + 
    #scaled_summonses + 
    #scaled_traffic_volume + 
    # Note: remove boro and nta because they are invariant within an intersection
    ns(scaled_pop_density, df = 2) +        
    year | 
    intersection_id,
  data = sites_ever_treated
)

summary(model_fixed)
```

```{r}
# exploring model_spline since it is the final chosen model

# Load broom (fixest works with broom automatically)
library(broom)

# Tidy the model output
# fixest automatically clusters SEs by the fixed effect (intersection_id)
irr_results_tidy <- tidy(model_fixed, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         se = "cluster") %>%   # Explicitly request clustered SEs
                   select(term, estimate, conf.low, conf.high, p.value)

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results
library(plotly)
library(dplyr)
library(fixest)

# --- FIX: Extraction method ---
# We extract from the summary table to ensure row counts match (ignoring theta)
# se = "cluster" ensures we use the clustered SEs you requested
model_summary <- summary(model_fixed, se = "cluster")$coeftable

# Extract columns by index to be safe (1=Est, 2=SE, 4=Pval)
coef_point_estimates <- model_summary[, 1]
model_se <- model_summary[, 2]
p_values_model <- model_summary[, 4]

# compute z-scores (Optional, since p-values are already extracted above, but good for checking)
z_scores <- coef_point_estimates / model_se

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"year",
  # "ns(scaled_pop_density, df = 2)1", 
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
# Using names(coef_point_estimates) works now because we extracted from the aligned table
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  # Filter to only include terms in the custom order
  filter(term %in% custom_order) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Fixed Effects Negative Binomial; SEs clustered by Intersection",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p
```




























