
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(data.table)
library(glmmTMB)
library(DHARMa)
library(effects)
library(performance)
library(purrr)
library(furrr)
library(splines)
library(future.apply)
library(broom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(plotly) 
library(ggeffects)
```

# Uploading and Preparing the Data

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

# Model - Year as a categorical variable

```{r}
# convert year to categorical
sites_ever_treated$year <- as.factor(sites_ever_treated$year)

# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    year +   
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Tidy the model output. This will automatically handle the NA coefficients.
irr_results_tidy <- tidy(model_spline, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         effects = "fixed") %>% # Only show fixed effects
                   select(term, estimate, conf.low, conf.high, p.value) # Optional: reorder

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# after introducing new terms, is there multicollinearity?

# very high

check_collinearity(model_spline)
```

```{r}
# diagnostics
# cannot do because of NA coefficients 
sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
```

# There is a big change in speed limit post (significantly negative -> significantly positive), slow zones (significantly negative -> non significant), and signal retiming (significantly positive -> non significant).











# Model - Year as a categorical variable, taking out all other time dependent effects

```{r}
# convert year to categorical
sites_ever_treated$year <- as.factor(sites_ever_treated$year)

# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    #covid_2020 + 
    #scaled_summonses + 
    #scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    year +   
    (1 | boro_name/ntaname/intersection_id),
  family = nbinom1,
  data = sites_ever_treated
) 

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Tidy the model output. This will automatically handle the NA coefficients.
irr_results_tidy <- tidy(model_spline, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         effects = "fixed") %>% # Only show fixed effects
                   select(term, estimate, conf.low, conf.high, p.value) # Optional: reorder

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# after removing terms, is there multicollinearity?

# very high for speed limit - this is probably what's causing the issue

check_collinearity(model_spline)
```

```{r}
# diagnostics
sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
AIC(model_spline)
BIC(model_spline)
```

# No imporvement for any diagnostic measure, if anythinng some got even more significant
# Worse AIC and BIC too

# There is a big change in speed limit post (significantly negative -> significantly positive), slow zones (significantly negative -> non significant), and signal retiming (significantly positive -> non significant).













# Model - Poisson version, all else held equal

```{r}
# uploading
sites_ever_treated <- read_csv('../data/output/intersection_intervention_table_ever_treated_2015-2022.csv', show_col_types = FALSE)
# meant to handle large datasets
sites_ever_treated <- as.data.table(sites_ever_treated)
```

```{r}
# scale variables
# the incidence ratio of these variables will now represent the change in odds for a *one standard deviation* increase in the original predictor
# nyc_metro_pop not included to to multicollnearity

sites_ever_treated <- sites_ever_treated %>%
  mutate(
    scaled_summonses = scale(citywide_traffic_summonses),
    scaled_traffic_volume = scale(nyc_metro_dvmt),
    scaled_pop_density = scale(pop_per_sqmi),
    scaled_time = scale(time)
  )
```

```{r}
# one spline for scaled_pop_density
model_spline <- glmmTMB(
  pedestrian_death_or_injury ~ 
    turn_traffic_calming_post +
    leading_pedestrian_interval_post +
    enhanced_crossing_post +
    speed_humps_post +
    street_improvement_project_post +
    street_improvement_corridors_post +
    signal_retiming_post +
    slow_zones_post +
    speed_limit_post +
    covid_2020 + 
    scaled_summonses + 
    scaled_traffic_volume + 
    ns(scaled_pop_density, df = 2) +        
    scaled_time +   
    (1 | boro_name/ntaname/intersection_id),
  family = poisson,
  data = sites_ever_treated
) 

summary(model_spline)
```

```{r}
# exploring model_spline since it is the final chosen model

# Tidy the model output. This will automatically handle the NA coefficients.
irr_results_tidy <- tidy(model_spline, 
                         conf.int = TRUE,      # Calculate confidence intervals
                         exponentiate = TRUE,  # Exponentiate to get IRRs
                         effects = "fixed") %>% # Only show fixed effects
                   select(term, estimate, conf.low, conf.high, p.value) # Optional: reorder

# Print the clean results table
print(irr_results_tidy)
```

```{r}
# visualizing results

# extract fixed effect estimates and standard errors 
coef_point_estimates <- fixef(model_spline)$cond
model_vcov <- vcov(model_spline)$cond
model_se <- sqrt(diag(model_vcov))

# compute z-scores and two-tailed p-values
z_scores <- coef_point_estimates / model_se
p_values_model <- 2 * (1 - pnorm(abs(z_scores)))

# compute confidence intervals (normal approximation)
ci_lower <- coef_point_estimates - 1.96 * model_se
ci_upper <- coef_point_estimates + 1.96 * model_se

# define custom order
custom_order <- c(
  #"scaled_time",
  # "ns(scaled_pop_density, df = 2)1", # can't be interpreted in the same way as the other coefficients
  # "ns(scaled_pop_density, df = 2)2",
  #"scaled_traffic_volume",
  #"scaled_summonses",
  #"covid_2020",
  "speed_limit_post",
  "slow_zones_post",
  "street_improvement_corridors_post",
  "street_improvement_project_post",
  "signal_retiming_post",
  "speed_humps_post",
  "enhanced_crossing_post",
  "leading_pedestrian_interval_post",
  "turn_traffic_calming_post"
)

# create percent change data frame
percent_change_df <- data.frame(
  term = names(coef_point_estimates),
  percent_change = 100 * (exp(coef_point_estimates) - 1),
  lower_ci = 100 * (exp(ci_lower) - 1),
  upper_ci = 100 * (exp(ci_upper) - 1),
  p_value = p_values_model
) %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    term = factor(term, levels = custom_order)
  )

# add significance indicator
percent_change_df <- percent_change_df %>%
  mutate(
    significant = ifelse(p_value < 0.05, "Significant", "Not Significant")
  )

# set factor levels for y-axis ordering
percent_change_df$term <- factor(percent_change_df$term, levels = custom_order)

# create interactive Plotly chart
p <- plot_ly(data = percent_change_df) %>%
  add_segments(
    x = ~lower_ci,
    xend = ~upper_ci,
    y = ~term,
    yend = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    showlegend = FALSE,
    hoverinfo = "none",
    line = list(width = 2)
  ) %>%
  add_markers(
    x = ~percent_change,
    y = ~term,
    color = ~significant,
    colors = c("Significant" = "blue", "Not Significant" = "red"),
    marker = list(size = 10),
    text = ~paste0(
      "Percent Change: ", round(percent_change, 2), "%<br>",
      "95% Bootstrap CI: (", round(lower_ci, 2), "%, ", round(upper_ci, 2), "%)<br>",
      "p-value: ", round(p_value, 4)
    ),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "Pedestrian Deaths and Injuries",
    subtitle = "Point estimates from model_dispformula; cluster-robust bootstrap SEs and CIs",
    xaxis = list(
      title = "Percent Change", 
      zeroline = TRUE, 
      zerolinecolor = 'black', 
      zerolinewidth = 1
    ),
    yaxis = list(
      title = "Fixed Effect", 
      categoryorder = "array", 
      categoryarray = rev(custom_order)
    ),
    legend = list(title = list(text = "Significance")),
    colorway = c("blue", "red")
  )

# show the plot
p

```

```{r}
# is there multicollinearity?

# no issues

check_collinearity(model_spline)
```

```{r}
# diagnostics

sim_res_spline <- simulateResiduals(fittedModel = model_spline, n = 500, plot = TRUE)
testZeroInflation(sim_res_spline) # zero-inflation test
testDispersion(sim_res_spline) # dispersion test
testUniformity(sim_res_spline) # KS test
testOutliers(sim_res_spline) # outliers
AIC(model_spline)
BIC(model_spline)
```

# Same exact results

















